---
title: "p8105_hw1_nk2814"
author: "Naama Kipperman"
date: "9/17/2019"
output: github_document
---



# Problem 1
### Creating a Data Frame and Taking means



```{r Create Dataframe and Mean}

# First, load tidyverse package.
library(tidyverse)   

# Set a seed to ensure reproducibility

set.seed(1)

# Next, define data frame called 'problem1_df'

problem1_df = tibble(    
  norm_samp = rnorm(8),    
  norm_samp_pos = norm_samp > 0,  
  vec_char = c("this", "is", "my", "first", "time", "using", "R", "software"),   
  vec_factor = factor(c("apple", "orange", "plum", "apple", "orange", "plum", "apple", "orange"))
)

# Try to take the mean of each variable in your dataframe. What works and what doesn’t?

mean(pull(problem1_df, norm_samp)) # mean is calculated. result below.
                                   

mean(pull(problem1_df, norm_samp_pos)) # mean is calculated. result below.
                                       
mean(pull(problem1_df, vec_char)) # unable to calculate the mean of this character vector. error message below.

mean(pull(problem1_df, vec_factor)) # unable to calculate the mean of this factor vector. error message below.

```

### Attempting to convert variables to numeric 

``` {r eval=FALSE} 
as.numeric(pull(problem1_df, norm_samp_pos))
as.numeric(pull(problem1_df, vec_char))
as.numeric(pull(problem1_df, vec_factor))

# The logical vector norm_samp_pos successfully converted to a numeric vector. 
# The character vector vec_char could not convert to a numeric vector - NA's were introduced by coercion instead. 
# This helps explain why we couldn't take the mean of this vector - it's a string of characters.
# The factor vector vec_factor successfully converted to a numeric vector of 1, 2, and 3. 

```

### Conversions of the logical vector
```{r logical vector conversions}

# Convert logical vector norm_samp_pos to numeric vector
# Multiply result by our random sample norm_samp

as.numeric(pull(problem1_df, norm_samp_pos))*pull(problem1_df, norm_samp) # Resulting vector below:


# Convert the logical vector norm_samp_pos to a factor 
# Multiply result by our random sample norm_samp
# Error message appears.
# Since we're working with a factor vector, it won't multiply our random sample of numbers

as.factor(pull(problem1_df, norm_samp_pos))*pull(problem1_df, norm_samp)



# Convert the logical vector norm_samp_pos to a factor 
# Conver the resulting factor variable to numeric
# Then, multiply the result by our random sample norm_samp

as.numeric(as.factor(pull(problem1_df, norm_samp_pos)))*pull(problem1_df, norm_samp) # Resulting vector below:



```

# Problem 2 
### Create a dataframe 

```{r Create dataframe}


# Next, define dataframe called problem2_df

problem2_df = tibble(    
  x = rnorm(500),
  y = rnorm(500),
  xy_logical = x+y> 1,  
  numeric_xy = as.numeric(xy_logical),
  factor_xy = as.factor(xy_logical),
)
```

### Short description of my dataframe

Number of rows in my dataframe: `r nrow(problem2_df)`.


Number of columns in my dataframe: `r ncol(problem2_df)`.


The mean of x is `r mean(pull(problem2_df, x))`.


The median of x is `r median(pull(problem2_df, x))`. 

The standard deviation of x is `r sd(pull(problem2_df, x))`. 

The proportion of cases where x+y > 1 is `r mean(pull(problem2_df, xy_logical))`. 

### Scatterplots

#### 1. scatterplot of y versus x, colored using xy_logical:
``` {r} 

ggplot(problem2_df, aes(x = x, y = y, color=xy_logical)) + geom_point()

ggsave("problem2_df.pdf")  # save first plot to current project

```

Since this scatterplot is colored using a binary logical vector, the color scale is binary(red and blue).


#### 2. scatterplot of y versus x, colored using numeric_xy:

```{r} 
ggplot(problem2_df, aes(x = x, y = y, color=numeric_xy)) + geom_point()

```

Since this scatterplot is colored using a numeric vector, which is technially continuous, the color scale is a color gradient.
However, since this vector only actually contains two possible values, there are only 2 colors seen on the plot.

#### 3. scatterplot of y versus x, colored using factor_xy

```{r} 
ggplot(problem2_df, aes(x = x, y = y, color=factor_xy)) + geom_point()

```

Since this scatterplot is colored using a factor vector that consists of two discrete factors, the scale is binary (red and blue). 
