---
title: "p8105_hw1_nk2814"
author: "Naama Kipperman"
date: "9/17/2019"
output: github_document
---



# Problem 1
### Creating a Data Frame and Taking means



```{r Create Dataframe and Mean}

# First, load tidyverse package.
library(tidyverse)   

# Set a seed to ensure reproducibility

set.seed(1)

# Next, define data frame called 'problem1_df'

problem1_df = tibble(    
  norm_samp = rnorm(8),    
  norm_samp_pos = norm_samp > 0,  
  vec_char = c("this", "is", "my", "first", "time", "using",   "R", "software"),   
  vec_factor = factor(c("apple", "orange", "plum", "apple", "orange", "plum", "apple", "orange"))
)

# Try to take the mean of each variable in your dataframe. What works and what doesn’t?

mean(pull(problem1_df, norm_samp)) # mean is calculated. result below.
                                   

mean(pull(problem1_df, norm_samp_pos)) # mean is calculated. result below.
                                       
mean(pull(problem1_df, vec_char)) # unable to calculate the mean of this character vector. error message below.

mean(pull(problem1_df, vec_factor)) # unable to calculate the mean of this factor vector. error message below.

```

### Attempting to convert variables to numeric 

``` {r eval=FALSE} 
as.numeric(pull(problem1_df, norm_samp_pos))
as.numeric(pull(problem1_df, vec_char))
as.numeric(pull(problem1_df, vec_factor))

# The logical vector norm_samp_pos successfully converted to a numeric vector. 
# The character vector vec_char could not convert to a numeric vector - NA's were introduced by coercion instead. This helps explain why we couldn't take the mean of this vector - it's a string of characters. R 
# The factor vector vec_factor successfully converted to a numeric vector of 1, 2, and 3. 

```

### Conversions of the logical vector
```{r logical vector conversions}

# Convert logical vector norm_samp_pos to numeric

logical_num=as.numeric(pull(problem1_df, norm_samp_pos))

# Print logical_num to make sure it successfully converted from 'True' and 'False' values to numeric values

view(logical_num)

# Problem 2 
